---
title: "Grammer of graphics-ggplot2"
author: "Swapnil Awasthi"
date: "10 November 2017"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Why ggplot2?
Very robust, useful and flexible for descriptive statistics. Advantages of ggplot2:

1-consistent, plot specification at a high level of abstraction. 

2-very flexible. 

3-theme system for polishing plot appearance. 

4-mature and complete graphics system. 

5-Great community support and active mailing list. 

NOTE: Data for this exercise can be downloaded from-http://tutorials.iq.harvard.edu/R/Rgraphics.zip

```{r data-prep}
library(ggplot2)
library(ggrepel)
housing <- read.csv('C:/Users/awasthi/Desktop/landdata-states.csv', header = TRUE, sep = ",",na.strings = c("NA", "#N/A","" ))
#checking null
colSums(is.na(housing))
#saving nulls in the df
nullhousing <- subset(housing, is.na(housing$region))
head(nullhousing)
#checking the State for which region is null
unique(nullhousing$State)
#all the NAs are from State = DC
subset(housing, housing$State == 'DC')[,2]
#setting DC to south region as per https://www2.census.gov/geo/pdfs/maps-data/maps/reference/us_regdiv.pdf
housing[housing$State=='DC',2] <- 'South'
#aggregating the data based home value - by state
Hvalue_by_state <- aggregate(Home.Value ~ State, housing, FUN = sum)
#creating a sorted dataset for further use
sortedd <- Hvalue_by_state[order(Hvalue_by_state$Home.Value, decreasing = TRUE),]
#top 5 states with highest home value
top5 <- as.list(sortedd[1:5,1])
#top 10 states with highest home value
top10states <- as.list(sortedd[1:10,1])
```

## Including Plots

```{r plot, echo=TRUE}
#scatter plot
ggplot(subset(housing, State %in% (unlist(top5))), aes(x=Date, y=Home.Value, color=State))+geom_point()
#box plot for top 10 State vs HomeValue #Box plots provide graphic display of five-number summary
ggplot(subset(housing, State %in% (unlist(top10states))), aes(x=State, y=Home.Value, color=State))+geom_boxplot()+geom_jitter(alpha=0.3, color='tomato')
#creating a dataframe based on a date/time-frame
hp.2007Q1 <- subset(housing, Date == '2007.25')

#let us take a closer look on scatter plot for 2007 1st Quarter - Land value vs Structure Cost
ggplot(hp.2007Q1, aes(x= Land.Value, y = Structure.Cost))+geom_point()
#since our x axis has a high variance - lets try to fit against a log transformation of Land Value
ggplot(hp.2007Q1, aes(x= log(Land.Value), y = Structure.Cost))+geom_point()
#color coding
ggplot(hp.2007Q1, aes(x= log(Land.Value), y = Structure.Cost, color = log(Land.Value)))+geom_point()

#Let us try to fit a line based on our X and Y and add it to your plot
hp.2007Q1$pred.SC <- predict(lm(Structure.Cost ~ log(Land.Value), data = hp.2007Q1))
p1 <- ggplot(hp.2007Q1, aes(x=log(Land.Value), y = Structure.Cost ))
#adding points and line
p1 + geom_point(aes(color = Home.Value))+geom_line(aes(y=pred.SC))
#mapping states on the plot
p1 + geom_text(aes(label = State), size = 2, color = 'red')
#install.packages("ggrepel") #library("ggrepel") required!!
#this is to map points with text
p1 + geom_point(color='red') + geom_text_repel(aes(label=State), size = 3)

#mapping variable to other aesthetics
p1 + geom_point(aes(color=Home.Value, shape = region))
#mapping variable to other aesthetics, adding one more dimension(size) to our plot
#here size is based on the home value - can be seen in the legends
p1 + geom_point(aes(size=Home.Value, shape = region, color = Home.Value))


#Statistical transformations
p2 <- ggplot(housing, aes(x=Home.Value))+geom_histogram(stat = 'bin', binwidth = 4000)
#top10states
top10 <- (sortedd[1:10,])
ggplot(top10, aes(x=State, y=Home.Value)) + geom_bar(stat = 'identity')
#all states included
ggplot(Hvalue_by_state, aes(x=State, y=Home.Value)) + geom_bar(stat = 'identity')

#Home value by Date and State
p3 <- ggplot(housing, aes(x=State, y=Home.Price.Index)) + theme(legend.position = 'top', axis.text = element_text(size = 6))
p3+geom_point(aes(color=Date), alpha=0.5,size=1.5)
#To avoid overlapping - we can use jitter
p4 <- p3+geom_point(aes(color=Date), alpha=0.5,size=1.5, position = position_jitter(width = 0.25, height = 0))
#color coding to indicate HIGHs and LOWs
p4+scale_color_continuous(low = 'green', high = 'red')

#Facets
p5 <- ggplot(housing, aes(x=Date, y=Home.Value))
p5 + geom_line(aes(color=State))  #Difficult to distinguish??? Facets can help
#Facets can help here #These can be used to see the correlation as well!
p5 + geom_line()+facet_wrap(~State, ncol = 10)
```

                                                  End of document
                                      Feedback/Suggestions are welcome, Thank you!
